Package Documentation

### 2022-06-22 quarto-doc/pkg_doc.qmd
Add quarto doc and its associated files. Re #97

---
title: "ITHIM"
author: "Rob Johnson, Ali Abbas"
date: "`r format(Sys.time(), '%d %B, %Y')`"

bibliography: ithim.bib

output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

@@

## Summary of literature

## Summary of aims

@@

The ITHIM as described here is implemented in the ITHIM-R R package (<https://github.com/ITHIM/ITHIM-R>). Other cities are also included in that package (São Paulo, Delhi, Bangalore), and reference will be made to those where relevant.

@@

!\[ITHIM workflow. Inputs to the model have a bold, black outline. The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)). To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box. \*AT: active travel. MET: metabolic equivalent task. pp: per person. AP: air pollution. PA: physical activity. DR: dose--response relationship. GBD: global burden of disease. BOD:

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("flowchart.pdf")
```

@@

::: center
::: {#indices}
| Label | Name                 | Levels              |
|:------|:---------------------|:--------------------|
| $a$   | Age group            | 15--49              |
|       |                      | 50--69              |
| $g$   | Gender               | Male                |
|       |                      | Female              |
| $h$   | Disease              |                     |
| $i$   | Individual index     |                     |
| $j$   | Trip index           |                     |
| $m$   | Transport mode       | walk                |
|       |                      | cycling (cyc.)      |
|       |                      | bus                 |
|       |                      | car                 |
|       |                      | motorbike (mot.)    |
|       |                      | goods vehicles (GV) |
|       |                      | subway (sub.)       |
| $o$   | Outcome              | Death               |
|       |                      | YLL                 |
| $s$   | Scenario             |                     |
| $w$   | AP RR hyperparameter | $1,2,3,4$           |
| $x$   | PA RR hyperparameter | $1,2,3$             |
| $z$   | PA dose              | $0,...,300$         |

: Indices used in the ITHIM variables. The "Label" is the subscript used to refer to the index in this document. The "Name" is descriptive. The "Levels" are the levels used in the Accra model.
:::
:::

::: center
::: {#abbrev}
| Abbreviation | Meaning                                       |
|:-------------|:----------------------------------------------|
| AP           | Air pollution                                 |
| CDF          | Cumulative distribution function              |
| COPD         | Chronic obstructive pulmonary disease         |
| DR           | Dose response                                 |
| EVPPI        | Expected value of perfect partial information |
| GAM          | Generalised additive model                    |
| GBD          | Global burden of disease                      |
| IHD          | Ischemic heart disease                        |
| ITHIM        | Integrated transport and health impact model  |
| LRI          | Lower respiratory infection                   |
| MMET         | Marginal metabolic equivalent task            |
| PA           | Physical activity                             |
| PAF          | Population attributable fraction              |
| PIF          | Potential impact fraction                     |
| PM2.5        | Particulate matter (diameter $< 2.5 \mu$m)    |
| PT           | Public transport                              |
| RR           | Relative risk                                 |
| VOI          | Value of information                          |
| YLL          | Years of life lost                            |

: Common abbreviations.
:::
:::

::: landscape
::: center
::: {#inputs}

@@

| -   Pollution model                       |                                                                 |                                    |                               |                    |
|                                           |                                                                 |                                    |                               |                    |
|     values\*                              |                                                                 |                                    |                               |                    |

@@

| -   Pollution & health model              |                                                                 |                                    |                               |                    |
|                                           |                                                                 |                                    |                               |                    |
|     values\*                              |                                                                 |                                    |                               |                    |

@@

: All values input to the ITHIM model.
:::
:::
:::

@@

From the synthetic population we create "scenarios" of different pictures of travel by changing certain trips (e.g. swapping the mode of travel for particular trips and/or particular people). For the purposes of illustration, we choose a simple scenario for Accra in which every person gains one one-kilometre walk.

@@

```{r out.height = "800px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("cancer_PA_dose_response.pdf")
```

@@

The XX uncertain univariate parameters are shown in Figure [3](#distributions){reference-type="ref" reference="distributions"}. Most of the parameters are self explanatory. Cycling and walking MMETs are the number of MMETs per hour when undertaking cycling and walking, and determine also the ventilation rates. Motorcycle distance is the total distance travelled by motorcycles relative to the total distance travelled by cars in the baseline scenario. Non-travel PA, injury reporting rate and NCD burden all act as scalars for the corresponding datasets. Note that the non-travel PA scalar does not affect the $\approx$<!-- -->40% of the population whose non-travel PA is 0.

```{r out.height = "460px", out.width='800px', fig.align='center', fg.cap = 'Distributions of univariate uncertain parameters (1024 samples).', echo=F}
knitr::include_graphics("uncertain_parameters.pdf")
```

@@

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("prob_zero_PA.pdf")
```

(){#prob_zero_PA width="50%"}

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
"Distributions of Dirichlet-distributed emission inventories for four modes and varying confidence values. Raw input values were 4, 4, 20, and 60."

knitr::include_graphics("emission_dist.pdf")
```

@@

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F, fig.show='hold', fig.cap="We use the same uniform samples to generate random variables from different distributions. It means that the samples are perfectly correlated."}
knitr::include_graphics("DR_transform.pdf")
```

@@

```{r out.height = "600px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("accra_walk_yll.pdf")
```

@@

```{r out.height = "460px", out.width='800px', fig.align='center', fig.cap="EVPPI for Accra",fig.path="accra_walk_evppi.pdf", echo=F}
#knitr::include_graphics("accra_walk_evppi.pdf")
```

@@

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("ithim_with_parameters.pdf")
```

@@

```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
knitr::include_graphics("PA_dose_response.pdf")
```

@@

::: center
::: ThreePartTable
::: TableNotes
CDF: cumulative distribution function. We use generative distributions to generate random numbers so that samples are correlated. For $t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$ with CDF CDF$_F$. (I'd like to find an improved notation for this.)
:::

::: {#calculationstab}
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| Name                                                                                                                                                         | Description                                                                                    | Label                                        |
+:=============================================================================================================================================================+:===============================================================================================+:=============================================+
| *Scenario generation*                                                                                                                                        |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| ...                                                                                                                                                          |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| *Travel calculations*                                                                                                                                        |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk} \}\, \forall\, T_j:m(j)=\text{bus}$                                                               | New walk trips to bus                                                                          | walk *to_bus_and* combine_scen               |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\bar{T}_{a,g,m,s}=\sum_{j:a(j )= a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$                                                                                        | (Weighted) total travel per age, gender, mode and scenario                                     |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| \$ \che ck{T}\_{a,g,m,s}=\frac{\bar{T}_{a,                                                                                                                   | Relative travel per mode and scenario by age and gender                                        | inj_distances\[\[1\]\]                       |
| g, m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}\$                                                                                                                      |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\hat{ T} _{m,s}=\sum_{a,g}\bar{T}_{a,g,m,s}$                                                                                                                | Total travel per mode and scenario                                                             | dist                                         |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{T}_{m,s}=\frac{\hat{T}_{ m, s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                      | Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1 | inj_distances\[\[2\]\]                       |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| *Injury calculations*                                                                                                                                        |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $E_{m_{\text{cas}}}=\psi\omega$                                                                                                                              | Casualty exponent                                                                              | CAS_EXPONENT                                 |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $E_{m_{\t ex t{str}}}=\omega-E_{m_{\text{cas}}}$                                                                                                             | Striker exponent                                                                               | STR_EXPONENT                                 |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}=\mathcal{S}\left(\sigma\c dot {I}_{a,g,m_{\text{vic}},m_{\text{| Fataliti                                   | es in scenarios by strike mode \| injuries                                                     |                                              |
| str}}},\bar{T}_{a,g,m,s},E_{m_{\t ex t{cas}}},E_{m_{\text{str}}}\right)$                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\check{I}_{a,g,m,o=\text{death                                                                                                                              | Fatalities in scenarios                                                                        | v ictim_deaths                               |
| },s} =\sum_{m_{\text{str}}}\tilde{I}_{ a, g,m_{\text{vic}},m_{\text{str}},s}$                                                                                |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $R_{a,g}=U_{a,g,h=\text{road                                                                                                                                 | Road YLL-to-death ratio                                                                        | GBD_INJ_YLL                                  |
| injuries},o=\text{YLL}}/U_{a,g,h= \t ext{road injuries},o=\text{death}}$                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\check{ I}_{ a,g,m,o=\text{YLL},s}=\check{I}_{ a, g,m,o=\text{death},s}\cdot R_{a,g}$                                                                       | YLLs scaled from fatalities                                                                    | deaths\_ yll_injuries                        |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| *Pollution calculations*                                                                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\dot{P}_{m}=\left\{\begin{array}{lr} \gam ma_m & P=1 \\ \tilde{\gamma}_m, \  | tilde{\gamma}_m \sim \text{Dir}(f(P,\  | gamma_m)) & P<1 \end{array}\right.$ | Emissions by mode \| \|                                                                        | trans_emissions                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{P}_{m,s}=\dot{P}_m\hat{T} _{ m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                     | Emissions by mode in scenario relative to baseline                                             | trans_emissions                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| \$ \h at{P}*{s}=*\sum{m} \tilde{P}\_{m,s}\$                                                                                                                  | Emissions scalar for scenarios                                                                 | baseline_sum                                 |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\bar {P}_s=\eta(\zeta\hat{P}_s + 1-\zeta)$ \|                                                                                                               | Background PM2.5 concentration in scenarios \|                                                 | scenario_pm                                  |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                 | On-road PM2.5 exposure ratio                                                                   | on_road_off_road_ratio                       |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{K}_{m\                                                                                                                                               | On-road PM2.5 exposure ratio for open vehicle                                                  | on_road_off_road_ratio                       |
| in\{\text{walk,cyc.,mot.}\},s}=K_s$                                                                                                                          |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\                                                                                                                                                           | Subway PM2.5 exposure ratio \|                                                                 | subway_ratio                                 |
| tilde{K}_{m\in\{\text{sub.}\},s}=C_6$ \|                                                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\                                                                                                                                                           | In-vehicle PM2.5 exposure ratio                                                                | in_vehicle_ratio                             |
| tilde{K}_{m\not\in\{\text{walk,cyc.                                                                                                                          |                                                                                                |                                              |
| ,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3)$                                                                                                                          |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $V_m=C_1+\frac{1}{2}\lambda_m$                                                                                                                               | Ventilation rates                                                                              | vent_rates                                   |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde {V}_{i,m,s} = \sum_{j:i(j)=i,m(j) =m ,s(j)=s}T_j\cdot V_{m(j)}/Z_{m(j)}$ \|                                                                          | In-travel air inhaled \|                                                                       |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\bar{V}_{i,s} = \sum_{m}\t il de{V}_{i,m,s}\cdot \tilde{K}_{m,s}$                                                                                           | In-travel PM2.5 inhalation rate                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| \$ \hat{V}*{i,s} = C_1*\left(24-\sum{j                                                                                                                       | Non-travel air inhaled                                                                         |                                              |
| :i(j)=i,s(j)=s}T_j/Z\_{m(j)}\right)\$                                                                                                                        |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| \$\check{W}*{i,s} =* \bar{ P}*s(*\bar{V}{i,s}+\hat{V}{i,s})/24                                                                                               | Total PM2.5 inhaled per hour                                                                   |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| %/(\sum\*m\                                                                                                                                                  |                                                                                                |                                              |
| \*\tilde{V}{i,m,s}+\hat{V}\_{i,s})\$                                                                                                                         |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ \|                                                                              | AP dose-- response--curve parameter \|                                                         | DR_AP_LIST                                   |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{H}_{a,h,w=3} =\te xt{CDF}_{H_{a,h,w=3}|\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                           | AP dose-- response--curve parameter                                                            | DR_AP_LIST                                   |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{H }_{a,h,w=3}=\text{CDF}_{H_{a,h, w=3} |\tilde{H}_{a,h,w=3},\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ \|                                   | AP dose-- response--curve parameter \|                                                         | DR_AP_LIST                                   |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{H}_{a,h,w=3}=\text{CD                                                                                                                                | AP dose-- response--curve parameter                                                            | DR_AP_LIST                                   |
| F}_{H_{a,h,w=3}|\tilde{H}_{a,h, w=3} ,\tilde{H}_{a,h,w=3},\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=3}\times$                                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\left\{1 - \exp\left(-\tilde{H }_ {a=a(i),h,w=3} (\check{W}_{i,s} - \                                                                                       | RR of disease given pollution                                                                  | RR_AP\_ calculations                         |
| tilde{H}_{a=a(i),h,w=3}) ^ {\tild e{ H}_{a=a(i),h,w=3} }\right)\right\}$                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| *Physical activity calculations*                                                                                                                             |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\dot{Y}_{a,g}=\                                                                                                                                             | Raw probability non-travel MMETs = 0 for demographic group $a,g$                               | raw_zero                                     |
| text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}$                                                                                                                          |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\b ar{Y}_{a,g}=\left\{\begin{array }{lr } \dot{Y}_{a,g} & \tilde{Y}=1 \\ \t ext{CDF}_{\text{Beta}(f(\tilde{Y},\                                             | Probability non-travel MMETs = 0 for demographic group $a,g$                                   | zeros                                        |
| dot{Y}_{a,g}))}^{-1}(\tilde{\thet a} ) & \tilde{Y}<1 \end{array}\right.$                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\check{Y}_i = \left\{\begin {a rray}{lr} 0 & \bar{Y}_{a(i),g(i)} \                                                                                          | Non-travel MMETs                                                                               | SYNTHETIC\_ POPULATION\$ work_l tpa_marg_met |
| \ \text{Sample}(\{Y_{i'}\}|_{a( i')= a(i),g(i')=g(i),Y_{i'}>0}) & 1-\b ar {Y}_{a(i),g(i)} \end{array}\right.$                                                |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $M_{i,s}=\sum_{j :i(j )=i,s(j)=s}(\chi T_j\cdot \lambda _{ m(j)}/Z_{m(j)})+\theta \check{Y}_i$                                                               | Total MMETs per person                                                                         | mmets                                        |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{G}_{h,x}(m) = {G}_{h, x,z= \lfloor{m}\rfloor}+({G}_{h,x,z=\l ce il{m}\rceil}-{G}_{h,x,z=\lfloor{m}\                                                  | Interpolated dose--response parameters                                                         | rr, lb, ub                                   |
| rfloor})\frac{m-\lfloor{m}\rfloor }{ \lceil{m}\rceil-\lfloor{m}\rfloor}$                                                                                     |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $F_S(s_x) = \mathcal{N}(s_{x=2},( s_ {x=3}-s_{x=1})/1.96)|_{0<F_S(s_x)}$                                                                                     | Truncated normal function                                                                      |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $W_{h,i,s}=\text{CDF}_{F_S\                                                                                                                                  | Relative risk of disease given PA                                                              | RR_PA\_ calculations                         |
| left (\tilde{G}_{h,x}(M_{i,s})\right)} ^{ -1}\left(\phi_{h}\right)^{\dagger}$                                                                                |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| *Burden-of-disease calculations*                                                                                                                             |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{W}_{h, i, s}=W_{h,i,s}\cdot\check{H}_{h,i,s}$                                                                                                        | Combined PA and AP relative risks                                                              | RR_PA_AP\_ calculations                      |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\hat{W}_{a,g,h,s}= \su m_ {i:a(i)=a,g(i)=g}\tilde{W}_{h,i,s}$                                                                                               | Populati on-attributable fractions                                                             | pif_temp                                     |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\bar{W}_{a,g,h,s}= (\hat{W} _{a, g,h,s=\text{scen1}}-\hat{W}_{a,g, h, s})/\hat{W}_{a,g,h,s=\text{scen1}}$                                                   | Populati on-attributable fractions relative to scenario 1                                      | pif_scen                                     |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\bar{U}_{a,g,h,o}=U_{a ,g ,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                               | GBD scaled to local population                                                                 | DISEASE_BURDEN                               |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\tilde{U}_{a, g, h,o}=\rho_h\cdot \bar{U}_{a,g,h,o}$                                                                                                        | Scaled background disease for Neoplasms, IHD, Lung cancer                                      | gbd_data_scaled                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\hat{U}_{a,g,h,o,s}=\bar{ W} _{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}$                                                                                            | Combined health and PIF                                                                        | yll_dfs, death_dfs                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
| $\hat{U}_{a,g,h=\text{road injury                                                                                                                            | Injury health burden                                                                           | inj                                          |
| },o,s}=\sum_m\check{I}_{a,g,m,o,s}$                                                                                                                          |                                                                                                |                                              |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+

: All calculations in the ITHIM.
:::
:::
:::
:::

# Uncertainty in travel data

We model uncertainty in travel via the definition of "travel attributes" (in the terminology of @Mohammadian2010). Specifically, we define the propensity to travel in a day, and the distance travelled in a day, given that travelled occurred. Each travel attribute is defined per demographic group, per mode.

In contrast to other travel-related synthetic populations, we have four demographic groups (rather than clusters) which consist of individuals (rather than households). A crucial difference between our objectives and the existing work on simulating populations in travel-demand models (that I've found so far) is that we don't need trip-level data for our calculations. Simulating trip-level data is computationally expensive and high-dimensional in terms of uncertain parameters. For our downstream computations, we need only summary data: travel per person per mode. Beyond scenario generation, there is no need for us to consider trips.

E.g. @Saadi2016 focuses on (a) synthetic population via matching covariates and (b) the sequence of trips taken by each person. (a) isn't a big thing for Accra; we're only using two covariates. And for (b), I think we can focus on people as it's a lot simpler. For our purposes, the things in this work that might be useful to us include matching principles for (a), when we revisit it, and distributions (in particular joint distributions) for travel. More useful to us would be distributions for total travel, but I haven't found any in this literature (e.g. zero-inflated compound Poisson distribution).

In contrast, my main questions are:

-   How to account for multi-mode trips.

-   What to do about zero travel for e.g. older cyclists. One possibility is to smooth over all modes and/or demographic groups.

-   In the ITHIM-R programme, are bus and truck drivers participants?

-   How should we include correlations between modes? This could also be aided by a smooth model across modes.

-   Whether there is a better method to get distance travelled per person than by resampling raw data.

## Method

### Summarise the travel survey

Our first step is to summarise the demographic information and the travel survey. In Table [5](#population){reference-type="ref" reference="population"} are the demographic data: the populations of the city, and the number of people in the survey. In Table [6](#calculations){reference-type="ref" reference="calculations"} are the summary statistics from the travel survey which we will use to generate synthetic travel. From the raw travel data (actually the adjusted trip set, which is the raw travel plus motorcycle, bus and truck trips), we calculate the probability to travel, $B_{a,g,m}$, as the number of people who travelled by that mode divided by the number of people in the demographic group. It is to these values that we assign uncertainty through the variable PROPENSITY_TO_TRAVEL.

::: center
::: {#population}
| Age    | Gender | Surveyed | Accra population |
|:-------|:-------|---------:|-----------------:|
| 15--49 | Male   |      279 |        5,847,716 |
| 50--69 | Male   |       56 |        1,016,476 |
| 15--49 | Female |      328 |        6,360,535 |
| 50--69 | Female |       69 |        1,110,037 |

: Population by demographic group.
:::
:::

::: center
::: ThreePartTable
::: TableNotes
:::

>>>>

---
title: "Package Documentation"
format:
  html:
    light: flatly
    dark: darkly
    toc: true

reference-location: margin
citation-location: margin

number-sections: true
crossref:
  chapters: true

---

```{r, echo=FALSE}
library(knitr)

io <- readRDS("../results/multi_city/io.rds")
cities <- names(io)[!names(io) %in% 'scen_prop']
scen_def <- io$scen_prop
```

@@

@@

The ITHIM as described here is implemented in the ITHIM-R R package (<https://github.com/ITHIM/ITHIM-R>). In the Latin America settings, we have a list of 20 cities, which are (Antofagasta, Arica, Belo Horizonte, Bogota, Buenos Aires, Cali, Copiapo, Coquimbo Laserena, Gran Valparaiso, Iquique Altohospicio, Medellin, Mexico City, Montevideo, Osorno, Puerto Montt, San Antonio, Santiago, Sao Paulo, Temuco Padrelascasas, Valdivia), and reference will be made to those where relevant.

@@

![ITHIM workflow. Inputs to the model have a bold, black outline. The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)). To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box. \*AT: active travel. MET: metabolic equivalent task. pp: per person. AP: air pollution. PA: physical activity. DR: dose--response relationship. GBD: global burden of disease. BOD: Burden of Disease](figures/flowchart.png "Flowchart: Model descrption"){fig-alt="A flowchart of model."}

@@

+---------------+----------------------+----------------------+
| Label         | Name                 | Levels               |
+===============+======================+======================+
| *a*           | Age group            | 15--49\              |
|               |                      | 50--69               |
+---------------+----------------------+----------------------+
| *g*           |                      | Male\                |
|               |                      | Female               |
+---------------+----------------------+----------------------+
| *h*           | Disease              |                      |
+---------------+----------------------+----------------------+
| *i*           | Individual index     |                      |
+---------------+----------------------+----------------------+
| *j*           | Trip index           |                      |
+---------------+----------------------+----------------------+
| *m*           | Transport mode       | walk\                |
|               |                      | cycling (cyc.)\      |
|               |                      | bus\                 |
|               |                      | car\                 |
|               |                      | motorbike (mot.)\    |
|               |                      | goods vehicles (GV)\ |
|               |                      | subway (sub.)        |
+---------------+----------------------+----------------------+
| *o*           | Outcome              | Death\               |
|               |                      | YLL                  |
+---------------+----------------------+----------------------+
| *s*           | Scenario             | 3                    |
+---------------+----------------------+----------------------+
| *w*           | AP RR hyperparameter | 1, 2, 3, 4           |
+---------------+----------------------+----------------------+
| *x*           | PA RR hyperparameter | 1, 2, 3              |
+---------------+----------------------+----------------------+
| *z*           | PA dose              |                      |
+---------------+----------------------+----------------------+

| Abbreviation | Meaning                                       |
|:-------------|:----------------------------------------------|
| AP           | Air pollution                                 |
| CDF          | Cumulative distribution function              |
| COPD         | Chronic obstructive pulmonary disease         |
| DR           | Dose response                                 |
| EVPPI        | Expected value of perfect partial information |
| GAM          | Generalised additive model                    |
| GBD          | Global burden of disease                      |
| IHD          | Ischemic heart disease                        |
| ITHIM        | Integrated transport and health impact model  |
| LRI          | Lower respiratory infection                   |
| MMET         | Marginal metabolic equivalent task            |
| PA           | Physical activity                             |
| PAF          | Population attributable fraction              |
| PIF          | Potential impact fraction                     |
| PM2.5        | Particulate matter (diameter $< 2.5 \mu$m)    |
| PT           | Public transport                              |
| RR           | Relative risk                                 |
| VOI          | Value of information                          |
| YLL          | Years of life lost                            |

: Indices used in the ITHIM variables. The "Label" is the subscript used to refer to the index in this document. The "Name" is descriptive. The "Levels" are the levels used in the Accra model.

```{r}

```

{#inputs}

@@

| *Pollution model values*                  |                                                                 |                                    |                               |                    |

@@

| *Pollution & health model values*         |                                                                 |                                    |                               |                    |

@@

@@

From the synthetic population we create "scenarios" of different pictures of travel by changing certain trips (e.g. swapping the mode of travel for particular trips and/or particular people). For the purposes of illustration, we have created three scenarios mentioned in the table @tbl-scen-specs, where each mode such as `cycle` has propensities in three different distance categories: `0-2 km`, `2-6 km` and `6+ km`. What the scenario does that in each of those distance categories, it increases the current mode share by the propensities. So for instance, if in a city the baseline mode share of `cycle` is `10` percent for short trips (`0-2 km`), then in the `cycle` scenario, it increases upto 15.6 (10 + 5.6).

```{r}
#| label: tbl-scen-specs
#| tbl-cap: Scenario propensities.
#| warning: false

kable(scen_def |> tibble::rownames_to_column("Scenarios"))
```

@@

@@

The XX uncertain univariate parameters are shown in Figure [3](#distributions){reference-type="ref" reference="distributions"}. Most of the parameters are self explanatory. Cycling and walking MMETs are the number of MMETs per hour when undertaking cycling and walking, and determine also the ventilation rates. Motorcycle distance is the total distance travelled by motorcycles relative to the total distance travelled by cars in the baseline scenario. Non-travel PA, injury reporting rate and NCD burden all act as scalars for the corresponding datasets. Note that the non-travel PA scalar does not affect the $\approx$\<40% of the population whose non-travel PA is 0.

@@

@@

@@

@@

@@

@@

@@

::: center
::: ThreePartTable
::: TableNotes
CDF: cumulative distribution function. We use generative distributions to generate random numbers so that samples are correlated. For $t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$ with CDF CDF$_F$. (I'd like to find an improved notation for this.)
